---
title: "R notes"
output:
  pdf_document: default
---

# R usage and notes

## Simple data structures

### Numbers

```{r}
# numbers
41
# big numbers and rationals turn to floats
10000000000000000000000000000000000 
2/3

```

### Strings


```{r} 
"A simple string"
```

## Interacting with Operating System

Directories and listing files:

```{r}
getwd()
setwd('tmp')
getwd()
list.files()
setwd('..')
```

## Toplevel Variables

Checking and dropping toplevel variables. 

```{r}
foo <- 42
bar <- "hello"
ls()
rm('foo')
ls()
```


## Getting Help

```{r}
find('getwd')
apropos('getwd')
example('getwd')
```


## Vectors

A vector can be defined with c(elements) or by using the range operator `:`.

```{r}
c(11, 22, 33, 44)
1:10
foo <- c(11, 22, 33, 44)
foo
```

Common functions for vectors:


```{r}
mean(foo)
range(foo)
fivenum(foo)
var(foo)
cor(foo, c(10, 19, 34, 42))
cumprod(foo)
cumsum(foo)
cummax(foo)
```

A negative index means all values apart from the listed one.
```{r}
foo
foo[-1]
foo[c(-1, -2, -3)]
```

Comparison of a vector to a scalar yields a bit vector. Positions of true values in a bit vector can be selected with `which`, and the resulting vector can be used to select the values.


```{r}
foo <- c(10, 4, 21, 9, 2, 24)
foo > 10
which(foo > 10)
foo[which(foo > 10)]
```


## Sets

BOOKMARK

Sorted vectors can be used as sets.

```{r}
sa = c("a", "b", "c", "d")
sb = c("b", "d", "e")
union(sa, sb)
intersect(sa, sb)
setdiff(sa, sb)
setequal(sa, sb)
```

The `%in%` operator can be used to make a membership bit vector.

```{r}
sa %in% sb
which(sa %in% sb)
```







